Supervised Reinforcement Learning with Partial States: \\ a Teaching Method for Action Selection for HRI

### 2017-07-17 paper.tex
rewriting and completting first parts

\maketitle
\begin{abstract}
Robots are expected to be pervasive in human society in the future

not possible to hardcode everything by hand

requirement for learning

hard to learn for HRI
\begin{itemize}
	\item real time
	\item cost of exploration (real humans)
	\item no clear reward policy and cannot request a human to reward everything
	\item need to make the most out of each sample
\end{itemize}

Partial state RL - accelerate learning

\end{abstract}
%===============================================================================

@@

Human-Robot Interaction (HRI) aims at having humans and robot living together
in the society, interacting socially in many different environments and
contexts. Robot are expected to behave appropriately regardless of the domain
of interaction. However, not all behaviours can be known in advance and can be
implemented in the robot before its deployment in the real world. Similarly to
 humans, robots need
to be able to learn new tasks, new actions policies, new social norms and how
to make sense of the world surrounding them.

When interacting in the real world, humans face huge spaces of almost unlimited
dimensions, and despite this complexity, adults and even more children are able
to learn quickly, often from only a few demonstrations or examples. Humans are
able to take the most out of every datapoint available to them. On the other
hand, recent impressive advances in Machine Learning such as Deep Learning~\cite{lecun2015deep}
have been made possible by the availability of huge dataset covering millions
of labeled examples or dozens of hours of recording. These progress have
targetting either perception, where labeled datapoints can be available or
action selection in virtual environments where the impact of actions is
limited.

Similarly to humans, robots interacting in the real world will have to be able
to learn quickly. One way of achieving this fast learning is to learn from
humans, i.e using human inputs to boostrap the learning. We argue that to be
capable to interact in the real world with humans, robots have to be able to
learn from different types of inputs: demonstrations, commands, rewards and 
also autonomously.

@@

The main framework to provide a robot with autonomous learning is Reinforcement
Learning (RL)~\cite{kober2013reinforcement,sutton1998reinforcement}. In RL, an
agent is interacting in an environment
providing positive or numerical reward in reaction to the agent's actions and
the agent learns an action policy to maximise the expected cummulative discounted
reward obtained by the agent.

Classic approaches RL relies on exploring virtual environments where the agent
tries to maximise the reward. In most of the cases where RL achieves success, it
had access to a virtual environment where the only real cost of exploring was
time and energy, and the agent can interact as long as needed to gather
enough information on the environment to find a correct action policy. When
these conditions are reunited, RL can achieve impressive results, as shown by
the victory of a computer in Backgammon back in the 90s~\cite{tesauro1995temporal} or more recently the success achieved in playing
Atari games based only on the image and the game rewards~\cite{mnih2015human}.

@@

In some cases, the state space is too big to rely only on random exploration to
reach a successful action policy. For example the game of Go in its larger board
size contains more than $10^{210}$ states, so exhaustive search is not possible
with today's ressources. Authors of \cite{silver2016mastering} started with
supervised learning from go master's games to learn a decent policy and then
used convolutional networks, self play and tree search to achieve super-human
capabilities and beatting humans best players.

@@

\subsubsection{Guiding the learning}
Rather than providing initial knowledge to the agents, humans can also guide the
agent throughout the learning by organising the tasks the robot will face, this
approach is known as scaffolding \cite{saunders2006teaching}. This allows to
teach the robot successive more complex policies by increasing the complexity of
the task presented and reusing previously learnt behaviours.

shaping
guidance

@@

As a way to enable a human supervisor to teach a robot an action policy while
interacting with its environment, we proposed the Supervised Progressively
Autonomous Robot Competencies (SPARC) in \cite{senft2015sparc}. SPARC is independent
of the learning algorithm used, but define an interaction dynamic between an
agent learning an action policy while interacting in an environment and a
supervisor teaching this agent an action policy. The main goal of SPARC is
enabling a robot to learn from a supervisor while ensuring that every action
executed by the robot has approval of the supervisor but without requiring this
supervisor to manually enforce every single action.

The main concept in SPARC is giving the supervisor the control over the agent
action and combine this with online learning on the agent side. This control is
achieved using a suggestion/correction mechanism whereas the agent is presenting
to the supervisor the action about to be executed and the supervisor can cancel
this action before its execution or not doing anything, letting the action being
executed after a delay. Additionally, the supervisor can select actions to be
executed by the agent. The actions selected or canceled by the supervisor can be
used as input with the current state by a learning algorithm to progressively
improve the future suggestions.

Giving power to an expert to correct the actions of the agent before the
execution can ensure that even in the early phases of the learning the agent can
have an appropriate action policy for the current environment as shown in Figure~\ref{fig:comparison}.

\begin{figure}
    \centering
    \includegraphics[width=0.9\linewidth]{./fig/motivation.pdf}
    \caption{Idealised expectation for performance, autonomy and workload for a
    Wizard of Oz, a autonomous learner and SPARC.}
    \label{fig:comparison}
\end{figure}

SPARC has been design with the idea of teaching a robot to interact with humans,
it is specially usefull when the pace of action selection is low (less than
1Hz), when the agent has to learn with a limited number of datapoints and when
undesired actions executed by an agent can have a high cost.

In \cite{senft2017supervised}, we presented a way to combine SPARC and RL, by
assigning a positive reward to every action executed by the robot as it would
have been passively or actvely validated by the supervisor. However this method
was not making use of any type of generalisation, it was directly mapping a
single state action pair to a reward, and as such has limited efficiency in
environments with a continuous space or a high dimensional and undeterministic
one.

>>>>

\maketitle
\begin{abstract}
As robots will integrate human society, they will have to satify humans'
expectations. However at the design time, it seems hardly possible to
know in advance exactly all the behaviours expected from the robot. As such to
be accepted in human society robots have to learn how to interact with humans.
But learning social interaction is a complex task. Firstly, no simulator of
human interactions precise enough exists today, so the learning will probably have
to be made online and in the real world.
This means that robots will have to learn directly from interacting with humans,
which limits the quantity of datapoint possible to gather and make exploration
riskier as actions will have real impacts on humans or the robot itself. Additionally,
no clear reward function exists for social interactions. As such we argue that
robots will highly profit from human expertise and guidance to learn social
interactions. And as the quantity of data is limited, they also will have to make the most out of each datapoint. We
propose to combine the Supervised Progressively Autonomous Robot Competencies
allowing a safer online learning and Reinforcement Learning based on partial
states rather than full states to accelerate the learning by using more
information form the human teacher.
\end{abstract}
%===============================================================================

@@

Human-Robot Interaction (HRI) aims at having humans and robot living together
in the society, interacting socially in many different environments and
contexts. Robot are expected to behave appropriately regardless of the domain
of interaction. However, not all behaviours can be known in advance and be
implemented in the robot before its deployment in the real world. Similarly to
 humans, robots need
to be able to learn new tasks, new actions policies, new social norms and how
to make sense of the world surrounding them.

When interacting in the real world, humans face huge spaces of almost unlimited
dimensions, and despite this complexity, humans, and especially children are able
to learn quickly, often from only a few demonstrations or examples. One of the
 reason could be that they are able to take the most out of every datapoint available to them. 
 On the other
hand, recent impressive advances in Machine Learning such as Deep Learning~\cite{lecun2015deep}
have been made possible by the availability of huge dataset covering millions
of labeled examples or dozens of hours of recording. These progress have
targetting either perception, where labeled datapoints can be available or
action selection in virtual environments where the impact of actions is
limited.

However, these immense labelled dataset or virtula worlds are not existent for
social interaction. To learn how to interact with humans, robots will have to
experiment in the real world facing real reactions from their actions. So they
have to learn quickly while consistently exhibiting a decent behaviours.
One way of ensuring this correct  behaviours is to use a human to supervised the
the robot during the early phase of learning, and this human can also provide
inputs to boostrap the learning, achieving a faster learning. This paper present
a novel method based on human supervision, feature highlighting and partial
state reinforcement learning to achieve this fast and safe learning.

@@

The main framework to provide a robot with autonomous learning is Reinforcement
Learning (RL)~\cite{kober2013reinforcement,sutton1998reinforcement}. In RL, an
agent is interacting in an environment
providing numerical reward in reaction to the agent's actions and
the agent learns an action policy to maximise the expected cummulative discounted
reward obtained by the agent.

Classic approaches RL generally rely on exploring virtual environments where the agent
tries to maximise the reward. In most of the cases where RL achieves success, it
had access to a virtual environment where the only real cost of exploring is
time and energy, and the agent can interact as long as needed to gather
enough information on the environment to find a correct action policy. When
these conditions are reunited, RL can achieve impressive results, as shown by
the victory of a computer in Backgammon back in the 90s~\cite{tesauro1995temporal} or more recently the success achieved in playing
Atari games based only on the image and the game rewards~\cite{mnih2015human}.

@@

In some cases, the state space is too big to rely only on random exploration to
reach a successful action policy. For example the game of Go in its larger board
size contains more than $10^{210}$ states, so exhaustive search is not possible
with today's ressources. Authors of \cite{silver2016mastering} started with
supervised learning from go master's games to learn a decent policy and then
used deep learning, self play and tree search to achieve super-human
capabilities and beatting humans best players.

@@

\subsubsection{Guiding the learning}
Rather than solely providing initial knowledge to the agents, humans can also guide the
agent throughout the learning in a more human-like teaching scenario.

A first possibility is organising the tasks the robot will face. 
This approach, known as scaffolding \cite{saunders2006teaching} progressively
increase the difficulty and complexity of the task or the environment as the
robot is learning to achieved complex policy not directly learnable in the final
environment.

Another possibility is shaping. For example, the TAMER framework
\cite{knox2009interactively}, allows to teach an agent an action policy in an
environment providing no rewards using a human observer to supply rewards to the
agent. One of the interest of this approach is that the teacher can adapt the
rewards to the current policy of the agent, for example starting to reward
positively suboptimal behaviours but which still improve the current policy and
later, as the agent improve its behaviour reward these same suboptimal behaviour
negatively and only the optimal ones positively.

A last possibility to guide the the learning is to bias the action selction.
In~\cite{thomaz2008teachable}, the authors propose to use a human supervisor to
supply  an agent both rewards and potential guidance or advise indicating where
the agent should pay attention, or what it should do next. And they show that
giving the power to the supervisor to bias the action selection can improve the
learning, making it faster, safer.

@@

As a way to enable a human supervisor to teach a robot an action policy while
interacting with its environment, we proposed the Supervised Progressively
Autonomous Robot Competencies (SPARC) in~\cite{senft2015sparc}. SPARC defines an
interaction dynamic between an agent learning an action policy while interacting
in an environment and a
supervisor teaching this agent what to do. The main goal of SPARC is
enabling a robot to learn from a supervisor while ensuring that every action
executed by the robot has approval of the supervisor but without requiring this
supervisor to manually enforce every single action.

The main concept in SPARC is giving the supervisor the control over the agent
action and combine this with online learning on the agent side. This control is
achieved using a suggestion/correction mechanism whereby the agent presents
to the supervisor the action about to be executed and the supervisor can cancel
this action before its execution or not doing anything, letting the action being
executed after a delay. Additionally, the supervisor can select actions to be
executed by the agent. The actions selected, accepted or canceled by the supervisor can be
used as input with the current state by a learning algorithm to progressively
improve the future suggestions thus reducing the reliance on the supervisors.

Giving power to an expert to correct the actions of the agent before the
execution can ensure that even in the early phases of the learning the agent can
have an appropriate action policy for the current environment as shown in
Figure~\ref{fig:comparison} while making the learning faster compared to
autonomous learning. (update figure with autonomous learning, classical learning from human rewards
and sparc)

\begin{figure}
    \centering
    \includegraphics[width=0.9\linewidth]{./fig/motivation.pdf}
    \caption{Idealised expectation for performance, autonomy and human workload for a
    an autonomous learner, an approach using human rewards and SPARC.}
    \label{fig:comparison}
\end{figure}

SPARC has been design with the idea of teaching a robot to interact with humans,
it is specially usefull when the frequency of action selection is low (less than
1Hz) so action can be evaluated by a human before being executed, when the agent
has to learn with a limited number of datapoints and when
undesired actions executed by an agent can have a high cost and are to be
avoided.

In \cite{senft2017supervised}, we presented a way to combine SPARC and RL, by
assigning a positive reward to every action executed by the robot as it would
have been passively or actvely validated by the supervisor. However this method
was not making use of any type of generalisation, it was directly mapping a
single state action pair to a reward, and as such had limited efficiency in
environments with a continuous or high dimensional space or undeterministic
transition from one state to another.
